---
title: "DATA 511 Project (R Notebook)"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
This project is for DATA 511 - Big Data Strategy for Samford University.We have 
been tasked to collect data on a topic we are curious about, compile data, and 
produce visualization to find trends in the data.

I am looking into email response times from my company, Jacksonville State.


I first needed to import all the packages needed for this project.
```{r include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
library(reshape)
```

Import the whole dataset from the project. 
```{r}
email_response <- read_csv("hw5_data.csv")
```

Calculating the correlation coefficient between my response time and the 
response time
```{r Overall Corr Coefficient, echo=TRUE}
email_response$mytime_num <- as.numeric(email_response$my_response_time)
email_response$responsetime_num <- 
  as.numeric(email_response$return_response_time)
response_corr <- cor(email_response$mytime_num, email_response$responsetime_num)
response_corr

```
The number above shows us that there is a very slight positive correlation bewteen my response time and the respondents repsonse time. This becomes very evident when looking a scatter plot with the data.
```{r}
email_viz01 <- ggplot(data=email_response, aes(x=my_response_time, 
                                             y=return_response_time)) +
  geom_point() +
  geom_smooth(method=lm,
              se=FALSE) +
  xlab('My Response Time') + 
  ylab('Return Response Time') +
  ggtitle("Email Response Time Scatter Plot")
  
email_viz01
```


We can break it down even further to examine if there was a stronger correlation between my response time while sending messages through Microdot Outlook or through Microsoft Teams. 


```{r Microsoft Teams Correlation Coefficient, echo=TRUE}
mt_messages <- email_response %>% filter(service == "MT")
mt_messages$mytime_num <- as.numeric(mt_messages$my_response_time)
mt_messages$responsetime_num <- 
  as.numeric(mt_messages$return_response_time)
mt_response_corr <- cor(mt_messages$mytime_num, mt_messages$responsetime_num)
mt_response_corr
```
We see from the number above that there is actually a slightly negative correlation in response time for conversations taking place through Microsoft Teams. 


```{r Microsoft Outlook Correlation Coefficient, echo=TRUE}
mo_messages <- email_response %>% filter(service == "MO")
mo_messages$mytime_num <- as.numeric(mo_messages$my_response_time)
mo_messages$responsetime_num <- 
  as.numeric(mo_messages$return_response_time)
mo_response_corr <- cor(mo_messages$mytime_num, mo_messages$responsetime_num)
mo_response_corr
```


Scatter plot for data showing differences in Microsoft Outlook and Microsoft 
Teams response times
```{r}
email_viz <- ggplot(data=email_response, aes(x=my_response_time, 
                                             y=return_response_time, 
                                             color=service)) +
  geom_point() +
  geom_smooth(method=lm,
              se=FALSE) +
  xlab('My Response Time') + 
  ylab('Return Response Time') +
  scale_colour_manual(values=c("#0072C6", "#7B83EB"),
                        name="Communication Platform",
                        breaks=c("MO", "MT"),
                        labels=c("Microsoft Outlook", "Microsoft Teams")) +
  theme(legend.position = "top")
  
email_viz
```


Importing data for the times emails were received
```{r}
received_times <- read_csv("email_times.csv")
```
Building out a histograms to show when people sent emails the most
```{r}
m <- mean(received_times$times)
std <- sqrt(var(received_times$times))
resp_viz <- ggplot(data = received_times, aes(x=times)) +
  geom_histogram() +
  xlab("Time of the Day") + 
  ylab("Number of Emails") +
  ggtitle("Time of Day Emails Were Received") + 
  theme(plot.title = element_text(lineheight=.8, face="bold"))
resp_viz
```

